Estimating Fixed-frame Galaxy Magnitudes in the SDSS Spectroscopic Survey

### 2002-02-15 docs/paper/paper.tex
writing
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@43 75dff457-8751-0410-ab1a-cd701654ce6b

\title{Modeling Galaxy Spectral Energy Distributions in the SDSS
Spectroscopic Survey}

@@

% Authorship determined by those I was directly involved with
% in performing this work as well as those responsible for the photo-z
% plates and those who have characterized the filter curves (since
% these might be publicly distributed). 
\author{
Michael R. Blanton\altaffilmark{\ref{NYU}},
%Tamas Budavari\altaffilmark{\ref{JHU}},
%Andrew J. Connolly\altaffilmark{\ref{Pitt}},
Istv\'an Csabai\altaffilmark{\ref{JHU}},
Mamoru Doi\altaffilmark{\ref{Tokyo}},
%Daniel Eisenstein\altaffilmark{\ref{Arizona}},
James E. Gunn\altaffilmark{\ref{Princeton}}, and
David W. Hogg\altaffilmark{\ref{NYU}}
%David J. Schlegel\altaffilmark{\ref{Princeton}}
%Julianne Dalcanton\altaffilmark{\ref{UW}},
%Jon Loveday\altaffilmark{\ref{Sussex}},
%Michael A. Strauss\altaffilmark{\ref{Princeton}},
%Mark SubbaRao\altaffilmark{\ref{Chicago}},
%David H. Weinberg\altaffilmark{\ref{Ohio}},
%John E. Anderson, Jr.\altaffilmark{\ref{Fermilab}},
%James Annis\altaffilmark{\ref{Fermilab}},
%Neta A. Bahcall\altaffilmark{\ref{Princeton}},
%Mariangela Bernardi\altaffilmark{\ref{Chicago}},
%J. Brinkmann\altaffilmark{\ref{APO}},
%Robert J. Brunner\altaffilmark{\ref{Caltech}},
%Scott Burles\altaffilmark{\ref{Fermilab}},
%Larry Carey\altaffilmark{\ref{UW}},
%Francisco J. Castander\altaffilmark{\ref{Chicago}, \ref{Pyrenees}},
%Andrew J. Connolly\altaffilmark{\ref{Pitt}},
%Istv\'an Csabai\altaffilmark{\ref{JHU}},
%Douglas Finkbeiner\altaffilmark{\ref{Berkeley}},
%Scott Friedman\altaffilmark{\ref{JHU}},
%Joshua A. Frieman\altaffilmark{\ref{Fermilab}},
%Masataka Fukugita\altaffilmark{\ref{CosmicRay},\ref{IAS}},
%G. S. Hennessy\altaffilmark{\ref{USNO}},
%Robert B. Hindsley\altaffilmark{\ref{USNO}},
%Takashi Ichikawa\altaffilmark{\ref{Tokyo}},
%\v{Z}eljko Ivezi\'{c}\altaffilmark{\ref{Princeton}},
%Stephen Kent\altaffilmark{\ref{Fermilab}},
%G. R.~Knapp\altaffilmark{\ref{Princeton}},
%D. Q.~Lamb\altaffilmark{\ref{Chicago}},
%R. French Leger\altaffilmark{\ref{UW}},
%Daniel C. Long\altaffilmark{\ref{APO}},
%Robert H. Lupton\altaffilmark{\ref{Princeton}},
%Timothy A.~McKay\altaffilmark{\ref{Michigan}},
%Avery Meiksin\altaffilmark{\ref{Edinburgh}},
%Aronne Merelli\altaffilmark{\ref{Caltech}},
%Jeffrey A. Munn\altaffilmark{\ref{USNO}},
%Vijay Narayanan\altaffilmark{\ref{Princeton}},
%Matt Newcomb\altaffilmark{\ref{CarnegieMellon}},
%R. C. Nichol\altaffilmark{\ref{CarnegieMellon}},
%Sadanori Okamura\altaffilmark{\ref{Tokyo}},
%Russell Owen\altaffilmark{\ref{UW}},
%Jeffrey R.~Pier\altaffilmark{\ref{USNO}},
%Adrian Pope\altaffilmark{\ref{JHU}},
%Marc Postman\altaffilmark{\ref{STScI}},
%Thomas Quinn\altaffilmark{\ref{UW}},
%Constance M. Rockosi\altaffilmark{\ref{Chicago}},
%Donald P. Schneider\altaffilmark{\ref{PennState}}, 
%Kazuhiro Shimasaku\altaffilmark{\ref{Tokyo}},
%Walter A. Siegmund\altaffilmark{\ref{UW}},
%Stephen Smee\altaffilmark{\ref{Maryland}},
%Yehuda Snir\altaffilmark{\ref{CarnegieMellon}},
%Chris Stoughton\altaffilmark{\ref{Fermilab}},
%Christopher Stubbs\altaffilmark{\ref{UW}},
%Alexander S.~Szalay\altaffilmark{\ref{JHU}},
%Gyula P.~Szokoly\altaffilmark{\ref{Potsdam}},
%Aniruddha R.~Thakar\altaffilmark{\ref{JHU}},
%Christy Tremonti\altaffilmark{\ref{JHU}},
%Douglas L. Tucker\altaffilmark{\ref{Fermilab}},
%Alan Uomoto\altaffilmark{\ref{JHU}},
%Dan vanden Berk\altaffilmark{\ref{Fermilab}},
%Michael S. Vogeley\altaffilmark{\ref{Drexel}},
%Patrick Waddell\altaffilmark{\ref{UW}},
%Brian Yanny\altaffilmark{\ref{Fermilab}},
%Naoki Yasuda\altaffilmark{\ref{NAOJ}},
%and Donald G.~York\altaffilmark{\ref{Chicago}}
}

@@

%% Mark off your abstract in the ``abstract'' environment. In the manuscript
%% style, abstract will output a Received/Accepted line after the
%% title and affiliation information. No date will appear since the author
%% does not have this information. The dates will be filled in by the
%% editorial office after submission.
\begin{abstract}
Broad-band measurements of galaxy flux suffer from the difficulty that
they probe different regions of the rest-frame galaxy spectrum for
galaxies at different redshifts.  Certain astronomical questions, such
as the evolution of the luminosity function of galaxies, require
transforming these inherently redshift-dependent measurements of
galaxy magnitudes into redshift-independent quantities. To prepare to
address these astronomical questions, investigated in detail in
subsequent papers, we implement here the method of \citet{csabai00a}
for fitting spectral energy distributions (SEDs) to broad-band
photometric observations, in the context of the optical observations
of the Sloan Digital Sky Survey (SDSS). Linear combinations of four
spectral templates can reproduce the five SDSS magnitudes of galaxies
to the precision of the photometry. Expressed in the correct
coordinate system, the distribution of the coefficients of the
templates is planar, and in fact nearly linear. The resulting
reconstructed SEDs (the sum of the templates) can be used to recover
fixed-frame magnitudes over a range of redshifts. We show that the
results of doing so yield consistent results within the SDSS Main
Galaxy Sample and the SDSS Luminous Red Galaxy Sample, in the sense
that within each sample the trend with redshift of intrinsic colors
for similar type galaxies are small. A potential problem in the
measurement of the SDSS filter response curves is noted. We also
validate the results using galaxy spectrophotometry.  All code and
templates from this paper are publicly available and well-documented;
they are designed such that they can easily be incorporated into other
applications. The code is easily adapted to handle galaxy observations
in any AB magnitude system in the rest-frame optical regime.
\end{abstract}

@@

In the future, we expect that observations of galaxies (and indeed of
any astronomical sources) will be performed using integral field
spectrographs or equivalent devices which combine high spatial
resolution and high spectral resolution. By that time, the phenomena
of broad-band filters and the quaint terminology surrounding their
usage --- magnitudes, $K$-corrections, color terms, {\it etc.} ---
will have long since been forgotten. However, today, virtually all
observations of galaxies are made through such filters, and special
care has to be taken to recover knowledge of galaxy spectral energy
distributions (SEDs) from these observations. This paper focuses on a
method of facing this problem which (in the optical wavelength regime)
is fast, robust, and consistent over a large range of redshifts.

This task is nontrivial because SEDs of galaxies contain important
information on scales considerably smaller than the width of typical
broad-band filters. It is further complicated by the fact that SEDs of
distant galaxies are redshifted due to the expansion of the
universe. Thus, the more distant the galaxy, the further the observed
bandpass is blueshifted relative to the rest-frame spectral energy
distribution of the object observed.

In the past, people have in general accounted for these effects using
``$K$-corrections'' applied to the observed magnitudes. That is, a
function $K(z)$ is added to the standard cosmological bolometric
distance modulus $\mathrm{DM}(z)$. Sometimes a single function has
been applied regardless of the galaxy type, though more recently it
has become standard to use a discrete set of $K(z)$ functions
depending on galaxy type (based either on morphology or spectral
features). As described in \cite{oke68a} and later papers, the
function is based on the projection of an assumed galaxy SED
redshifted to $z$ onto the measured instrument response as a function
of wavelength (including the effects of the atmosphere, the
reflectivity of the mirrors, the filter transmission, and the response
of the CCD device or photographic emulsion).

However, galaxies do not all have the same SED, nor are they selected
from some discrete set of SEDs. For this reason, schemes for applying
the $K$-corrections can fail to be self-consistent: the SED assumed
for the $K$-corrections can be significantly inconsistent with the
observed galaxy colors! As we demand more precision from our
astronomical data analysis in new, large multi-band surveys, and in
particular as we try to quantify the evolution of galaxies, we must
take a more sophisticated approach to approximating fixed-frame
observations of galaxies.

Our approach here is to use a method for inferring the underlying SEDs
of a set of galaxies at a range of redshifts by requiring that their
SEDs all be drawn from a similar population. For each galaxy, we will
recover a model SED, which can be used to synthesize the galaxy's
magnitude in any bandpass. In addition, this method provides a natural
way to infer an estimate of the flux within the whole optical range.
The approach is equivalent (nearly identical) to the photometric
metric redshift estimation methods of \citet{csabai99a} and
\citet{budavari00a}, except we use slightly different coordinate
systems.

We implemented this system and are publishing it in order to prepare
the reader for upcoming papers which will rely heavily on the
reliability of the SED fits described here.  However, having
implemented the fits, there is no reason not to share them with the
community; thus, this paper serves the secondary purpose of describing
a piece of software being released to the astronomical community. The
computer software created is a C shared object library, around which
we have written both stand-alone C programs and IDL routines. We have
made the source code available publicly, through a URL on the World
Wide Web and through a public CVS repository. Improvements or ports to
other languages implemented by users may be incorporated into the code
upon request. The conditions of use for the code are that this paper
is cited in any resulting refereed journal article and that the
version of the code used is specified in any such paper. The version
of the code used to make the figures for this paper is {\tt kcorrect
v1\_2}.

Section \ref{sedfit} describes our method of fitting galaxy SEDs to
broad-band photometry. Section \ref{data} applies the method to
galaxies in the SDSS, showing that the fits are robust. Section
\ref{kcorrection} shows how one derives $K$-corrections from the
results, and discusses how best to use the results of such
fits. Section \ref{conclusion} concludes and discusses future
development of the method described here.

\section{Fitting SEDs to Galaxy Broad-band Colors}
\label{sedfit}

@@

Let us begin with an SED space defined by $N_b$ template galaxy SEDs
$\vv{v}_i(\lambda)$ (say from \citealt{bruzual93a}), where $N_b$ is
large. In principle, this should be a complete set of galaxy SEDs if
we want to express {\it any} galaxy SED in this basis set; in
practice, this property is not necessary.  Given a set of
$\vv{v}_i(\lambda)$, one can find an orthonormal set of spectra
$\vv{b}_i(\lambda)$ spanning the space.  To be more precise, we define
the dot-product in the space of galaxy spectra to be:
\begin{equation}
\vv{v}_i \cdot \vv{v}_j =
\int_{\lambda_{\mathrm{min}}}^{\lambda_{\mathrm{max}}} d\lambda
v_i(\lambda) v_j(\lambda),
\end{equation}
where $\lambda_{\mathrm{min}}$ and $\lambda_{\mathrm{max}}$ define the
wavelength range used to define orthogonality. These limits are
defined not over the full range covered by the $\vv{v}_i(\lambda)$
spectra, but instead over a smaller range which is within observable
wavelengths over most of the redshift range of the sample. If we
defined orthogonality over the full range of $\vv{v}(\lambda)$, much
of which is unobservable, it would complicate the interpretation of
our fits.  This choice does not affect the results for the
reconstructed SED fit quantitatively whatsoever.\footnote{If the
template spectra were identical in the subrange chosen and only
differed outside of it, this choice would cause computational
problems, but in fact the spectra are all independent in the
restricted, as well as the full, wavelength range.}  We then define
the basis of the SED space such that the $N_b$ template SEDs,
$\vv{v}_i$, are linear combinations of $N_b$ basis spectra $\vv{b}_i$
and that $\vv{b}_i \cdot \vv{b}_j = \delta_{ij}$. These conditions do
not fully specify the $\vv{b}_i$, which naturally may be rotated
arbitrarily within our $N_b$-dimensional subspace of spectral space.

@@

We define the ``total flux'' $f_k$ of each template as the flux in the
range $\lambda_{\mathrm{min}}<\lambda<\lambda_{\mathrm{max}}$; then an
estimate of the flux in this range is $F_{{t}}=\Sigma_k a_k f_k$. The
locus of points of a fixed flux in the component space $a_k$ lie on a
plane. Therefore, a natural choice for the orientation of the axes
$\vv{e}_k$ is to let $\vv{e}_0$ be perpendicular to the planes of
constant flux; in this manner, the coefficient $a_0$ in Equation
(\ref{lincomb}) is directly proportional to $F$ and is thus purely a
measure of the object's flux (in the wavelength range considered),
while the parameters $a_i/a_0$ (where $i>0$) primarily measure the
shape of the galaxy SED in the wavelength range considered. $F$ can be
trivially related to $L$, the luminosity in this range, by the inverse
square of the luminosity distance.  As it turns out, the parameters
$a_i/a_0$ tend to be distributed very approximately in an ellipsoid,
with not much curvature within the space defined by the template
SEDs. Thus, we rotate the axes $\vv{e}_i$ (for $i>0$) such that they
are aligned with the principal axes of the ellipsoid. Furthermore, we
normalize the vectors $\vv{e}_i$ such that the zeroth component
$a_0=F_t$ is in units of ergs cm$^{-2}$ s$^{-1}$. This choice of
coordinate system is arbitrary and does not affect our quantitative
results, but it is convenient.

In this way, we can reconstruct SEDs based on the broad-band
magnitudes of galaxies. From these reconstructed SEDs one can estimate
$K$-corrections, develop a measure of galaxy spectral type, or to
synthesize a very broad-band measure of the galaxy flux.  The
templates determined during the procedure can, of course, be used to
estimate photometric redshifts, as \citet{csabai00a} and
\citet{budavari00a} describe.

The version of the $K$-correction code ({\tt v1\_2}) implementing the
calculations described here, along with eigentemplates and filter
curves is publicly available on the World Wide Web at {\tt
http://physics.nyu.edu/\~\ mb144/kcorrect}. In addition, a public CVS
repository can be accessed at {\tt xxx} (instructions for exporting
the appropriate version of the code from this repository are provided
on the web page noted above. The whole of the code can be used through
the Research Systems, Incorporated, IDL language; everything except
for the template-fitting also exists in stand-alone C programs (which
call the same routines, guaranteeing consistency).

@@

In this section we describe the SDSS data set, we describe how we fit
the templates and the coefficients in detail for this data set, and we
describe our distribution of code to do this.

>>>>

\title{Estimating Fixed-frame Galaxy Magnitudes in the SDSS
Spectroscopic Survey}

@@

% Authorship determined by those I was directly involved with
% in performing this work as well as those responsible for the photo-z
% plates and those who have characterized the filter curves (since
% these might be publicly distributed). 
\author{
Michael R. Blanton\altaffilmark{\ref{NYU}},
%Tamas Budavari\altaffilmark{\ref{JHU}},
%Andrew J. Connolly\altaffilmark{\ref{Pitt}},
Istv\'an Csabai\altaffilmark{\ref{JHU}},
Mamoru Doi\altaffilmark{\ref{Tokyo}},
Daniel Eisenstein\altaffilmark{\ref{Arizona}},
James E. Gunn\altaffilmark{\ref{Princeton}}, and
David W. Hogg\altaffilmark{\ref{NYU}}
%David J. Schlegel\altaffilmark{\ref{Princeton}}
%Julianne Dalcanton\altaffilmark{\ref{UW}},
%Jon Loveday\altaffilmark{\ref{Sussex}},
%Michael A. Strauss\altaffilmark{\ref{Princeton}},
%Mark SubbaRao\altaffilmark{\ref{Chicago}},
%David H. Weinberg\altaffilmark{\ref{Ohio}},
%John E. Anderson, Jr.\altaffilmark{\ref{Fermilab}},
%James Annis\altaffilmark{\ref{Fermilab}},
%Neta A. Bahcall\altaffilmark{\ref{Princeton}},
%Mariangela Bernardi\altaffilmark{\ref{Chicago}},
%J. Brinkmann\altaffilmark{\ref{APO}},
%Robert J. Brunner\altaffilmark{\ref{Caltech}},
%Scott Burles\altaffilmark{\ref{Fermilab}},
%Larry Carey\altaffilmark{\ref{UW}},
%Francisco J. Castander\altaffilmark{\ref{Chicago}, \ref{Pyrenees}},
%Andrew J. Connolly\altaffilmark{\ref{Pitt}},
%Istv\'an Csabai\altaffilmark{\ref{JHU}},
%Douglas Finkbeiner\altaffilmark{\ref{Berkeley}},
%Scott Friedman\altaffilmark{\ref{JHU}},
%Joshua A. Frieman\altaffilmark{\ref{Fermilab}},
%Masataka Fukugita\altaffilmark{\ref{CosmicRay},\ref{IAS}},
%G. S. Hennessy\altaffilmark{\ref{USNO}},
%Robert B. Hindsley\altaffilmark{\ref{USNO}},
%Takashi Ichikawa\altaffilmark{\ref{Tokyo}},
%\v{Z}eljko Ivezi\'{c}\altaffilmark{\ref{Princeton}},
%Stephen Kent\altaffilmark{\ref{Fermilab}},
%G. R.~Knapp\altaffilmark{\ref{Princeton}},
%D. Q.~Lamb\altaffilmark{\ref{Chicago}},
%R. French Leger\altaffilmark{\ref{UW}},
%Daniel C. Long\altaffilmark{\ref{APO}},
%Robert H. Lupton\altaffilmark{\ref{Princeton}},
%Timothy A.~McKay\altaffilmark{\ref{Michigan}},
%Avery Meiksin\altaffilmark{\ref{Edinburgh}},
%Aronne Merelli\altaffilmark{\ref{Caltech}},
%Jeffrey A. Munn\altaffilmark{\ref{USNO}},
%Vijay Narayanan\altaffilmark{\ref{Princeton}},
%Matt Newcomb\altaffilmark{\ref{CarnegieMellon}},
%R. C. Nichol\altaffilmark{\ref{CarnegieMellon}},
%Sadanori Okamura\altaffilmark{\ref{Tokyo}},
%Russell Owen\altaffilmark{\ref{UW}},
%Jeffrey R.~Pier\altaffilmark{\ref{USNO}},
%Adrian Pope\altaffilmark{\ref{JHU}},
%Marc Postman\altaffilmark{\ref{STScI}},
%Thomas Quinn\altaffilmark{\ref{UW}},
%Constance M. Rockosi\altaffilmark{\ref{Chicago}},
%Donald P. Schneider\altaffilmark{\ref{PennState}}, 
%Kazuhiro Shimasaku\altaffilmark{\ref{Tokyo}},
%Walter A. Siegmund\altaffilmark{\ref{UW}},
%Stephen Smee\altaffilmark{\ref{Maryland}},
%Yehuda Snir\altaffilmark{\ref{CarnegieMellon}},
%Chris Stoughton\altaffilmark{\ref{Fermilab}},
%Christopher Stubbs\altaffilmark{\ref{UW}},
%Alexander S.~Szalay\altaffilmark{\ref{JHU}},
%Gyula P.~Szokoly\altaffilmark{\ref{Potsdam}},
%Aniruddha R.~Thakar\altaffilmark{\ref{JHU}},
%Christy Tremonti\altaffilmark{\ref{JHU}},
%Douglas L. Tucker\altaffilmark{\ref{Fermilab}},
%Alan Uomoto\altaffilmark{\ref{JHU}},
%Dan vanden Berk\altaffilmark{\ref{Fermilab}},
%Michael S. Vogeley\altaffilmark{\ref{Drexel}},
%Patrick Waddell\altaffilmark{\ref{UW}},
%Brian Yanny\altaffilmark{\ref{Fermilab}},
%Naoki Yasuda\altaffilmark{\ref{NAOJ}},
%and Donald G.~York\altaffilmark{\ref{Chicago}}
}

@@

%% Mark off your abstract in the ``abstract'' environment. In the manuscript
%% style, abstract will output a Received/Accepted line after the
%% title and affiliation information. No date will appear since the author
%% does not have this information. The dates will be filled in by the
%% editorial office after submission.
\begin{abstract}
Broad-band measurements of galaxy fluxes suffer from the difficulty
that for galaxies at different redshifts the observations probe
different regions of the rest-frame galaxy spectrum. Certain
astronomical questions, such as the evolution of the luminosity
function of galaxies, require transforming these inherently
redshift-dependent magnitudes into redshift-independent quantities. To
prepare to address these astronomical questions, investigated in
detail in subsequent papers, we implement here the method of
\citet{csabai00a} for fitting spectral energy distributions (SEDs) to
broad-band photometric observations, in the context of the optical
observations of the Sloan Digital Sky Survey (SDSS). Linear
combinations of four spectral templates can reproduce the five SDSS
magnitudes of all galaxies to the precision of the
photometry. Expressed in the appropriate coordinate system, the
distribution of the coefficients multiplying the templates is planar,
and in fact nearly linear. The resulting reconstructed SEDs can be
used to recover fixed-frame magnitudes over a range of redshifts. We
show that the results of doing so yield consistent results, in the
sense that within each sample the intrinsic colors of similar type
galaxies are nearly constant with redshift. We note one remaining
problem associated with the gap between the $g$ and $r$ bands. We
compare our results to simpler interpolation methods and galaxy
spectrophotometry.  The software which generates these results is
publicly available and easily adapted to handle galaxy observations in
any AB magnitude system in the rest-frame optical regime.
\end{abstract}

@@

In the future, we expect that observations of galaxies (and indeed of
any astronomical sources) will be performed using integral field
spectrographs or equivalent devices which combine high spatial
resolution and high spectral resolution. By that time, the phenomena
of broad-band filters and the quaint terminology surrounding their
usage --- magnitudes, $K$-corrections, color terms, {\it etc.} ---
will have long since been forgotten. However, today, virtually all
observations of galaxies are made through such filters, and special
care has to be taken to recover knowledge of galaxy spectral energy
distributions (SEDs) from these observations.  Reconstructing galaxy
SEDs is nontrivial because SEDs of galaxies contain important
information on scales considerably smaller than the width of typical
broad-band filters. Furthermore, SEDs of distant galaxies are
redshifted such that the more distant the galaxy, the further the
observed bandpass is blueshifted relative to the rest-frame spectral
energy distribution of the object observed.  This paper focuses on a
method of facing these problems which (in the optical wavelength
regime) is fast, robust, and consistent over a large range of
redshifts.

In the past, people have in general accounted for these effects using
``$K$-corrections'' applied to the observed magnitudes. In these
analyses, a function $K(z)$ is added to the standard cosmological
bolometric distance modulus $\mathrm{DM}(z)$. Sometimes a single
function has been applied regardless of the galaxy type, though more
recently it has become standard to use a discrete set of $K(z)$
functions depending on galaxy type (based either on morphology or
spectral features). As described in \cite{oke68a} and later papers,
the function is based on the projection of an assumed galaxy SED
redshifted to $z$ onto the measured instrument response as a function
of wavelength (including the effects of the atmosphere, the
reflectivity of the mirrors, the filter transmission, and the response
of the CCD device or photographic emulsion).  However, galaxies do not
all have the same SED, nor are they selected from some discrete set of
SEDs. For this reason, these schemes for applying the $K$-corrections
can fail to be self-consistent: the SED assumed for the
$K$-corrections can be significantly inconsistent with the observed
galaxy colors! As we demand more precision from our astronomical data
analysis in new, large multi-band surveys, and in particular as we try
to quantify the evolution of galaxies, we must take a more
sophisticated approach to approximating fixed-frame observations of
galaxies.

Our approach here is to use a method for inferring the underlying SEDs
of a set of galaxies at a range of redshifts by requiring that their
SEDs all be drawn from a similar population. For each galaxy, we will
recover a model SED, which can be used to synthesize the galaxy's
magnitude in any bandpass. Although the operation we are performing on
the magnitudes is not strictly speaking a ``$K$-correction,'' we will
refer to it as such in this and subseqent papers. Our approach is
equivalent (nearly identical) to the photometric metric redshift
estimation methods of \citet{csabai00a} and \citet{budavari00a},
except we use slightly different coordinate systems. In fact, the
software includes a fast and relatively accurate photometric redshift
estimator based on their method.

We implemented this system and are publishing it in order to prepare
the reader for upcoming papers which will rely heavily on the
reliability of the fixed-frame magnitudes determined here.  In
addition, having implemented the fits, there is no reason not to share
them with the astronomical community; thus, this paper serves the
secondary purpose of describing the release of a piece of
software. The computer software we distribute builds itself into a C
shared object library, around which we have written both stand-alone C
programs and IDL routines. We have made the source code available
publicly, through a URL on the World Wide Web and through a public CVS
repository. Improvements or ports to other languages implemented by
users may be incorporated into the code upon request. The conditions
of use for the code are that this paper is cited in any resulting
refereed journal article and that the version of the code used is
specified in any such paper. The version of the code used to make the
figures for this paper is {\tt kcorrect v1\_2}.

Section \ref{sedfit} describes our method of fitting galaxy SEDs to
broad-band photometry. Section \ref{data} applies the method to
galaxies in the SDSS, showing that the fits are robust. Section
\ref{kcorrection} shows how one derives $K$-corrections from the
results, and discusses how best to use the results of such
fits. Section \ref{conclusions} concludes and discusses future
development of the method described here.

\section{Calculating Fixed-frame Galaxy Magnitudes}
\label{sedfit}

First, we describe how we reconstruct galaxy SEDs from broad-band
magnitude measurements. Second, we describe how we convert these SEDs
into estimates of fixed-frame galaxy magnitudes.

\subsection{Fitting SEDs to Galaxy Broad-band Magnitudes}

@@

Let us begin with an SED space defined by $N_b$ template galaxy SEDs
$\vv{v}_i(\lambda)$ (say from \citealt{bruzual93a}), where $N_b$ is
large. In principle, this should be a complete set of galaxy SEDs if
we want {\it all} galaxy SEDs to be in the space spanned by the
$\vv{v}_i(\lambda)$; in practice, this property is not necessary.
Given a set of $\vv{v}_i(\lambda)$, one can find an orthonormal set of
spectra $\vv{b}_i(\lambda)$ spanning the same space.  To be more
precise, we define the dot-product in the space of galaxy spectra to
be:
\begin{equation}
\vv{v}_i \cdot \vv{v}_j =
\int_{\lambda_{\mathrm{min}}}^{\lambda_{\mathrm{max}}} d\lambda
v_i(\lambda) v_j(\lambda),
\end{equation}
where $\lambda_{\mathrm{min}}$ and $\lambda_{\mathrm{max}}$ define the
wavelength range used to define orthogonality. These limits are
defined not over the full range covered by the $\vv{v}_i(\lambda)$
spectra, but instead over a smaller range which is within observable
wavelengths over most of the redshift range of the
sample. \footnote{Were we to define orthogonality over the full range
of $\vv{v}(\lambda)$, much of which is unobservable, the geometry of
the distribution of galaxies in this space would be more complicated.
If the template spectra were identical in the subrange chosen and only
differed outside of it, our choice would cause computational
problems, but in fact the spectra are all independent in the
restricted, as well as the full, wavelength range.}  We then define
the basis of the SED space such that the $N_b$ template SEDs,
$\vv{v}_i$, are linear combinations of $N_b$ basis spectra $\vv{b}_i$
and that $\vv{b}_i \cdot \vv{b}_j = \delta_{ij}$. These conditions do
not fully specify the $\vv{b}_i$, which naturally may be rotated
arbitrarily within our $N_b$-dimensional subspace of spectral space.

@@

We define the ``total flux'' $f_k$ of each template as the flux in the
range $\lambda_{\mathrm{min}}<\lambda<\lambda_{\mathrm{max}}$; then an
estimate of the flux in this range is $F_{{t}}=\Sigma_k a_k f_k$.
Galaxies of some fixed $F_t$ clearly lie on a plane in the component
space $a_k$. Therefore, a natural choice for the orientation of the
axes $\vv{e}_k$ is to let $\vv{e}_0$ be perpendicular to the planes of
constant flux; in this manner, the coefficient $a_0$ in Equation
(\ref{lincomb}) is directly proportional to $F_t$ and is thus purely a
measure of the object's flux (in the wavelength range
$\lambda_{\mathrm{min}} <\lambda < \lambda_{\mathrm{max}}$), while the
parameters $a_i/a_0$ (where $i>0$) primarily measure the shape of the
galaxy SED in the wavelength range $\lambda_{\mathrm{min}} <\lambda <
\lambda_{\mathrm{max}}$. $F_t$ can be trivially related to $L_t$, the
total luminosity in this range, by the inverse square of the
luminosity distance (from, for example, \citealt{hogg99a}).  As it
turns out, the parameters $a_i/a_0$ tend to be distributed
approximately in an ellipsoid, with not much curvature within the
space defined by the template SEDs. Thus, we rotate the axes
$\vv{e}_i$ (for $i>0$) such that they are aligned with the principal
axes of the ellipsoid. Furthermore, we normalize the vectors
$\vv{e}_i$ such that the zeroth component $a_0=F_t$ is output by the
code in units of ergs cm$^{-2}$ s$^{-1}$. This choice of coordinate
system is arbitrary and does not affect our quantitative results, but
it is convenient.

In this way, we can reconstruct SEDs based on the broad-band
magnitudes of galaxies. From these reconstructed SEDs one can estimate
$K$-corrections, develop a measure of galaxy spectral type, or
synthesize other measurements of galaxy flux.  The templates
determined during the procedure can, of course, be used to estimate
photometric redshifts, as \citet{csabai00a} and \citet{budavari00a}
describe.

We here define a notation which we will use in subsequent papers for
the effective rest-frame bandpass at any redshift in band $b$, namely:
\begin{equation}
\band{z}{b}
\end{equation}
read, ``$b$-band at redshift $z$.'' The goal of ``$K$-corrections'' is
to convert the $ugriz$ bands at redshift $z$ for each galaxy to
$ugriz$ bands at some fixed redshift $z_0$.

Given AB galaxy magnitudes $^{z}m_{AB}$ and errors for a galaxy at
redshift $z$, one can use the method of the previous subsection to
reconstruct the galaxy rest-frame SED.  With this reconstruction in
hand, it is straightforward to project the SED onto bands at any
redshift $z_0$ ({\it cf.} Equation \ref{ABdef}):
\begin{eqnarray}
^{z_0}m_{\mathrm{rec}} &=& -2.41 - 2.5 \log_{10}\left[
\frac{\int_{0}^{\infty} d\lambda \frac{\lambda}{1+z_0}
f_{\mathrm{rec}}(\frac{\lambda}{1+z_0}) R(\lambda)}
{\int_{0}^{\infty} d\lambda \lambda^{-1} R(\lambda)}\right] \cr
&=& -48.60 - 2.5 \log_{10}\left[
\frac{\int_{0}^{\infty} d\nu \nu^{-1} (1+z_0)
f_{\mathrm{rec}}[\nu (1+z_0)] R(\nu)}
{\int_{0}^{\infty} d\nu \nu^{-1} R(\nu)}\right] 
\end{eqnarray}
The simple way to think about this equation is that simply moving away
from the Earth doing does not change the bolometric flux of the
galaxy. So when the spectrum is stretched by the $(1+z_0)$ redshift
factor, as in $f(\lambda/(1+z_0))$, the bolometric flux should be
constant; that is, you must apply the extra multiplicative factor
$1/(1+z_0)$ to counteract the stretching of the spectrum.

Given the reconstructed magnitudes, one can calculate the fixed-frame
magnitudes in one of two ways:
\begin{enumerate}
\item Simply use $\band{z_0}{m_{\mathrm{rec}}}$, the reconstructed
magnitudes at redshift $z_0$, as the fixed-frame magnitudes.
\item Use the difference
$\band{z_0}{m_{\mathrm{rec}}}-\band{z}{m_{\mathrm{rec}}}$ as an
estimate of the correction to apply to the observed $\band{z}{m}$.
\end{enumerate}
If the reconstructions $\band{z}{m_{\mathrm{rec}}}$ are very similar
to the actual observations $\band{z}{m}$ (as we will show is the case
for our fits), there is very little difference between these two
methods. Having estimated $\band{z_0}{m}$, you can then calculate the
luminosity by simply applying to the calculated magnitude the
cosmological distance modulus (that is, the luminosity
distance-squared law, as tabulated by, {\it e.g.}  \citealt{hogg97a}).

We want to emphasize here that, while $K$-correcting to a fixed frame
bandpass is {\it sometimes} necessary in order to achieve a scientific
objective, it is not {\it always} necessary or appropriate. Because
$K$-corrections are an inherently uncertain business (the broad-band
magnitudes just do not uniquely determine the SED) they should be
avoid or minimized when possible. For example, if one was calculating
the evolution of clustering of red and blue galaxies separately, it
would perhaps be wise to perform the separation not on the
$K$-corrected colors of the galaxies but on the median color of, say,
$M_\ast$ galaxies as a function of redshift. Similarly, in situation
where $K$-corrections cannot be ignored, such as the calculation of
the evolution of the luminosity function, their effect should be
minimized by, for example, correcting to the median redshift of
galaxies in the sample.

Now, typically one's measurements in will be difficult to connect to,
say, solar luminosities in $\band{z}{b}$ (unless $z=0$), simply
because people have not projected the appropriate stellar
spectrophotometry onto these bandpasses. However, it is our position
that stars are better understood than galaxies, and that it therefore
is in the end simpler to stay as close as possible to the system in
which the galaxies are observed. In any case, astronomy is quickly
reaching a level of precision for which the exact nature of the
bandpasses used has to be known and considered in most analyses of
observational data.

The reader may ask why, if we are basing our $K$-corrections on a full
model of the galaxy SED, we do not correct to bandpasses with simpler
shapes (say, Gaussians). The reason is that we want the effective
bandpasses to actually have been observed for some galaxies in the
sample, so that the $K$-correction for those galaxies is formally and
actually zero. This property is desirable because, as noted above,
there are uncertainties in the $K$-corrections. Nevertheless, we note
that synthesizing such magnitudes may be appropriate in certain
situations, having the virtue (similarly to the $AB$ magnitude system)
that such magnitudes are very easy to comprehend and synthesize from
theory. For example, if we wanted to convert our results into a form
easily comprehended by the future astronomers discussed in the first
paragraph of Section \ref{motivation}, it would be best to calculate
fluxes within wide top-hat filters so that the difficulties of
understanding the magnitude system were minimized.

\subsection{Public Access to the Code}

The version of the $K$-correction code ({\tt v1\_2}) implementing the
calculations described here, along with eigentemplates and filter
curves is publicly available on the World Wide Web at {\tt
http://physics.nyu.edu/\~\ mb144/kcorrect}. In addition, a public CVS
repository can be accessed at {\tt xxx} (instructions for exporting
the appropriate version of the code from this repository are provided
on the web page noted above). The whole of the code can be used
through the Research Systems, Incorporated, IDL language; everything
except for the template-fitting also exists in stand-alone C programs
(which call the same routines, guaranteeing consistency).

@@

In this section we describe how we applied the above method to the
SDSS data set.
